---
title: "HW6"
author: "Sam Mottahedi"
date: "August 3, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('~/Projects/STAT502/HW6/')
```

# Problem 1)
```{r}
data <- read.table('pearls.txt', header=T, colClasses = c('numeric', 'factor','factor','factor'))

mu <- mean(data$value)
mu_i <- tapply(data$value, data$no_coats, mean)
mu_j <- tapply(data$value, data$batch, mean)

tau_b <- mu - mu_j
tau_a <- mu - mu_i
```


## (a) 

$Y_{ijk} = \mu_{..} + \alpha_i + \beta_j + (\alpha \beta)_{ij} + \epsilon_{ijk}$

the terms $\alpha_i, (\alpha \beta)_{ij}, \epsilon_{ijk}$ are random and $\mu_{..}, \beta_j$ are fixed.


## (b)

|Mean sq| df | Expected MS| 
|-------|-------------|-----------------------------|
|MSA| $a-1=2$| $\sigma^2 + n \sigma_{\alpha \beta}^2 + \frac{bn}{a-1} \sum_j \alpha_i^2$|
|MSB| $b-1=3$|$\sigma^2 + an \sigma_{\beta}^2$|
|MSAB| $(a-1)(b-1)=6$| $\sigma^2 + n \sigma_{\alpha \beta}^2$|
|MSE| $ab(n-1)=36$| $\sigma^2$| 

$\sigma^2 = 4.832$

$\sigma_{\alpha \beta}^2 = `r (0.309 - 4.823)/4`$

$\sigma_{\beta} = `r (50.951-4.823)/(3* 4)`$

$E\{MSAB\} = `r 4.823 + 4*(0)`$

$E\{MSB\} = `r 4.823 + 3* 4* 3.844`$

$E\{MSA\} = `r 4.823 + 4 * 0 + (3*4/2) * sum(tau_b^2)`$


## (c)

```{r}
fit2 <- lm(value ~ no_coats + batch + no_coats:batch, data)
anova(fit2)
```



$H_0: \sigma^2_{\alpha \beta} = 0$

$H_a: \sigma^2_{\alpha \beta} > 0$


$F* = \frac{MSAB}{MSE}= `r 0.309/4.923` < F = `r qf(0.95, 6, 36)`$

concluding null, interaction term is not significant.


## d)

$H_0: \sigma^2_{\alpha} = 0$

$H_a: \sigma^2_{\alpha} > 0$

$F^* = \frac{75.194}{0.309} = `r 75.194/0.309`$

$F = `r qf(0.95, 2, 2*3)`$

$F^* > F$ concluding $H_a$ the effect of coat factor is significant.


## e)

|Mean sq| df | Expected MS| 
|-------|-------------|-----------------------------|
|MSA| $a-1=2$|$\sigma^2 +  bn \sigma_{\alpha}^2$|
|MSB| $b-1=3$|$\sigma^2 + n \sigma_{\alpha \beta}^2 + \frac{an}{b-1} \sum_j \beta_j^2$|
|MSAB| $(a-1)(b-1)=6$| $\sigma^2 + n \sigma_{\alpha \beta}^2$|
|MSE| $ab(n-1)=36$| $\sigma^2$| 

$\sigma^2 = 4.832$

$\sigma_{\alpha \beta}^2 = `r (0.309 - 4.823)/4`$

$\sigma_{\alpha} = `r (75.194-4.823)/(4* 4)`$

$E\{MSAB\} = `r 4.823 + 4*(0)`$

$E\{MSA\} = `r 4.823 + 4* 4* 4.3981875 `$

$E\{MSB\} = `r 4.823 + 4 * 0 + (4*4/3) * sum(tau_b^2)`$


## f)

```{r}
mu = tapply(data$value, data$no_coat, mean)
l1 = mu[2] - mu[1]
l2 = mu[3] - mu[1]
s = sqrt((0.309/(4*4))*(2))
t = qt(1- 0.1/(2*2),  3*4*3)
```


$B = `r qt(1- 0.1/(2*2),  3*4*3)`$

$s\{ \hat{D} \} = `r sqrt((0.309/(4*4))*(2))`$

$\mu_6 - \mu_8 = `r l1`$

$\mu_6 - \mu_{10} = `r l2`$

$`r l1 - s*t` \le \mu_8 - \mu_6 \le `r l1 + s*t`$

$`r l2 - s*t` \le \mu_8 - \mu_{10} \le `r l2 + s*t`$


# Problem 2)


### a) 

$Y_{ijk} = \mu_{..}   + \alpha_i + \beta_{j(i)} + \epsilon_{ijk}$

$\mu_{..} constant variable$

$\alpha_{i}: \textrm{constant related to factor A and subject to} \sum \alpha_i=0$

$\beta_{j(i)}: random normal variable with mean 0$

$\\epsilon_{ijk} independet N(0, \sigma^2)$

$i=1, \dots, 2; j=1, \dots, 3; k=1, \dots, 5$


### b)

$SSA = 0.01825$

$SSB(A) = 0.01153 + 0.44249 =  0.45402$

$SSE = 0.29020$

|         Source| SS    | df       | MS             | $E\{ MS \}$                                                   |
|---------------|-------|----------|----------------|---------------------------------------------------------------|
|Factor A       |0.01825|a-1=1     |$`r 0.01825 /1`$  |$\sigma^2 + bn \frac{\sum \alpha_i^2}{a-1} + n \sigma_{\beta}^2$|
|Factor B(in A) |0.45402|a(b-1)=4  |$`r 0.45402 /4`$  |$\sigma^2 + n \sigma_{\beta}^2$                                |
|Error          |0.29020|ab(n-1)=24|$`r 0.29020 /24`$ |$\sigma^2$                                                     |
|Total          |0.47227|abn-1=29  |$`r 0.47227 /29`$ | |


assuming $\alpha = 0.05$

for factor A: 

$F^* = MSA/MSB(A) = `r 0.01825/(0.45402 /4)`$

$F = `r qf(0.95,1, 4)`$

$F^* < F$ 

concluding $H_0$, there's no considerable variability in different sites.

for factor B: 

$F^* = MSB(A)/MSE = `r (0.45402 /4)/(0.29020 /24)`$

$F = `r qf(0.95,4, 24)`$

$F^* > F$  concluding $H_a$, there is considerable batch to batch difference.


# Problem 3)

```{r}
data = read.table('floral.dat',header=T,sep='\t') 
trt = as.factor(data$trt)
plot = as.factor(data$plot)
time = as.factor(data$time) 
y = data$resp 
```


## a)

```{r}
interaction.plot(time, trt, y)
```

The mean of y for unburned region is higher compared to the burned region. The mean of Y increases with time for unburned region. For the burned region the 
the mean of y increases with time between 1-9 and deacreases between 9-12 and again increases between 12-15.



## b)

$Y_{ijk} = \mu_{...} + \rho_{i(j)} + \alpha_{j} = \beta_k + (\alpha \beta)_{jk} + \epsilon_{ijk}$

$\mu_{...}$ is constant

$\rho_{i(j)}$ random ~ $N(0, \sigma_p^2)$, nested with in factor A.

$\alpha_j$ burn/no burn, fixed

$\beta_k$ : Time factor, fixed

## c)

That the covariance for the same plot has compound symmetry and all covariances should be similar in magnitude. Yes.

## d)

Yes, except for the two extreme points in left and right the rest of the points seems to be normal.


```{r}
fit <- lm(y ~ trt/plot + trt + time + trt:time)
anova(fit)
hist(fit$residuals, 20)
qqnorm(fit$residuals)
```


## e)

$H_0: all \alpha_j = 0$
$H_a: \textrm{not all} \alpha_j \textrm{equal zero}$

$F^* = \frac{MSA}{MSS(A)} = `r 9964/222.8`$

$F = `r qf(0.95, 1, 2*(9-1))`$

$F^* > F$ we conclude $H_a$. That treatment effect exist and burned areas are significantly different.

