---
title: "HW1"
author: "Sam Mottahedi"
date: "July 1, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd('~/Projects/STAT502/HW1/')
set.seed(1234)
require(ggplot2)
require(stargazer)
require(nlme)
```


(adapted from KNNL 16.8) The data in “paper.txt” are from an experiment investigating the effect of paper color on response rates for a questionnaire. 


(a) Create a box plot of the data separating the results by color. Does it appear there are differences in the response rates?  Yes.

```{r}
df <- read.table('./data/paper.txt', header = F)
df <- df[, c(1,2)]
df$V2 <- as.factor(df$V2)
p <- ggplot(df, aes(x= (V2), y=V1)) + geom_boxplot()

print(p)
```

(b) Calculate the mean and standard deviation of the response rate for each of the three colors. Also calculate the pooled sample standard deviation. 

```{r}
df.mean <- tapply(df$V1,df$V2,mean) 
df.sd <- tapply(df$V1,df$V2,sd) 
df.var <- tapply(df$V1,df$V2,var) 

pooled.sd <- sqrt(sum(4 * tapply(df$V1,df$V2, var)) / (15 - 3))
pooled.sd
```



(c) Denoting the population means of the blue and green groups by 1 and 2, respectively, carry out the test of $H_0 : \mu_1 - \mu_2 = 0$ versus the two-sided alternative. Report the degrees of freedom, test statistic, and conclusion with $\alpha = 0.05$. 

```{r, echo=T}
pooled.sd <- sqrt(sum((5-1) * df.var[1:2]) / (5 + 5 - 2))
t.value <- (df.mean[1] - df.mean[2])/ (pooled.sd * sqrt(1/5 + 1/5))
qt(0.025, 8)


```

$df = 8$

$t-value = -0.0911$

$p-value = 2 \times pt(0.0911) = `r 2* pt(-0.0911, 8)$

whe can't reject the null hypothesis. 

(d) Estimate 1 2 with 95% confidence. 


```{r, eval=T}
df.mean[1] - df.mean[2] - abs(qt(0.025, 8)) * pooled.sd * sqrt(1/5 + 1/5)
df.mean[1] - df.mean[2] + abs(qt(0.025, 8)) * pooled.sd * sqrt(1/5 + 1/5)
```
$lb =  -5.262716$  


$ub = 4.862716$

(e) What assumptions are you making for parts (c) and (d)? 


Two the samples have approximately the same variance.



# Problem 2

```{r}
data=read.table("./data/CH01PR19.txt")
y=data[,1]
x=data[,2]
fit=lm(y~x) 
summary(fit) 
```
## a)

$y = 2.11405 + 0.03883 \times x$


## b)

The estimated standard deviation of the coefficient. 

## c)

$\beta_{lb} = \beta  - t_{\alpha /2,n-2} \times Std. Error = `r 0.03883 - qt(0.95, 118) *  0.01277`$

$\beta_{up} = \beta + t_{\alpha /2,n-2} \times Std. Error = `r 0.03883 + qt(0.95, 118) *  0.01277`$


## d)

$y = 2.11405 + 0.03883 \times x = `r 2.11405 + 0.03883 * 30`$


## e)

```{r}
plot(x, y, xlab='ACT', ylab='GPA')
abline(fit)
```


# Problem 3 


## a)

```{r, echo=TRUE}
df1 <- data[data$V2 <= 25,]
df2 <- data[data$V2 > 25,]
fit1 <- lm(V1 ~ V2, data=df1)
fit2 <- lm(V1 ~ V2, data=df2)
```

$mean(GPA_low) = `r mean(df1$V1)`$

$mean(GPA_high) = `r mean(df2$V1)`$

$mean difference = `r mean(df2$V1) - mean(df1$V1)`$


## b)


```{r}
x1 = 1 * (x>25)
```

The $x > 25$ returs True and False. When multiplied by 1 the returned value is 0 and 1.

## c)

```{r, echo=FALSE}
fit3 <- (lm(y ~ x1))
summary(fit3)
```

$difference expected GPA = 2.94951$

$Std error = 0.07845$

$df = 120 -2 = 118$


## d)

$t = \frac{\beta}{Std. Error} = `r 0.27173/ 0.11588`$

$t^* = `r abs(qt(0.025, 118))`  < 2.345$

$P-value = `r 2 * (1- pt(2.345, 118))`$


## e)

Yes.

```{r}
low=y[x<=25];high=y[x>25] 

```

Returns the value of y if the conidtion in bracket is true. 


## f)


$mean_{high} = `r mean(high)`$

$mean_{low} = `r mean(low)`$


$differece_{mean} = `r mean(high) -mean(low)`$


$df = 118$ is equal to df in part c.

## g)

Yes, because both d, and g tests the difference in mean in two groups.

```{r}
t.test(low, high, var.equal = T)
```





