---
title: "HW2"
author: "Sam Mottahedi"
date: "July 8, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Problem 1 

Source| SS | DF | MS | F | p-value
------:|----|----|----|---|--------
Between treatments| 64.42| 3| $`r 64.42/3`$ | 8.98| $`r 1 - pf(8.98, 3, 21-4)`$
Within treatments|$`r 2.39 * (21-4)`$ |$`r 21-4`$ | 2.39 | |
Total| $`r 64.42 +2.39 * (21-4)`$ | 20| | |




# Problem 2

```{r}
y <- c(3129, 3000 ,2865 ,2890,3200, 3300 ,2975 ,
       3150,2800 ,2900 ,2985 ,3050,2600 ,2700, 
       2600, 2765) 
x <- rep(1:4, times=1, each=4)

fit <- aov(y ~  as.factor(x))
summary(fit)
```


Source| SS | DF | MS | F | p-value
------:|----|----|----|---|--------
Between treatments| 489740| 3| 163247| 12.73|0.000489
Within treatments|153908|12 | 12826 | |
Total|$`r 153908 + 489740`$| 15| | |

$p-value < 0.05$ The mean tensile strength for the four different mixing techniques are different.









# Problem 3


## a)

```{r, echo=FALSE }
library(psych)
df <- read.table('./CH16PR08.txt')[, 1:2]
df$V2 <- as.factor(df$V2)
# fit <- lm(V1 ~ -1 + V2, df)
# summary(fit)
```

```{r}
tapply(df$V1, df$V2, mean)
tapply(df$V1, df$V2, var)
var(df$V1)
describeBy(df$V1, df$V2)
```

## b)


$se_{blue} = 1.63$

$se_{green} = 1.47$

$se_{orange} = 1$


## c)

```{r}
summary( aov(V1 ~ V2, df))
```

$p-value = 0.684 > 0.05$

We can't rejecting the null ($H_0: \mu_1 = \dots  = \mu_4$). The mean differences are not significant.


## d)

```{r}
c1 <- c(1, 0, -1)
c2 <- c(0, 1, -2)
mat <- cbind(c1, c2)
contrasts(df$V2) <- mat

fit <- aov(V1 ~ V2, df)
summary.aov(fit, split=list(V2=list('orange vs blue'=1, 'orange vs green'=2)))
```


since both p-values for "orange vs. blue" and "orange vs green" are greater than $0.05$ we can't reject the null hypothesis $H_0:\mu_{orange} = \mu_{green}$ and $H_0:\mu_{orange} = \mu_{blue}$.




# Problem 4

## a)



Source| SS | DF | MS | F | p-value
------:|----|----|----|---|--------
Drug|56.10| 3| $`r 56.10/3`$ |$`r (56.10/3)/(62.12/24)`$ | $`r 1 - pf((56.10/3)/(62.12/24), 3, 24)`$
Error|62.12|24 | $`r 62.12/24`$ | |
Total|$`r 153908 + 489740`$| 27| | |


$`r (56.10/3)/(62.12/24)`$ | $`r 1 - pf((56.10/3)/(62.12/24), 3, 24)` < 0.05$

rejecting the null hypothesis $H_0:\mu_1 = \dots = \mu_4$, there is significant difference between bio-activity levels of the four drugs.


## b) 

Six.


## c)

```{r}
qt(1-0.05/2, 28-4)
```

## d)


Tukey:

```{r}
qtukey(0.95, 4, 24)/sqrt(2)
```

benferroni:

```{r}
qt(1-0.95/(2*6), 24)
```

scheffe:

```{r}
S2 = (4-1) * qf(0.05, 4-1, 28-4)
sqrt(S2)
```


## e)

```{r}
u1 = 66.1; u2 = 65.75; u3= 62.63; u4 = 63.85

l1 = u1 - u2; l2= u1 - u3
l3 = u1 - u4; l4 = u2- u3
l5= u2-u4; l6=u4-u3

mse = 2.5883333


s2 = mse*(1/7 + 1/7)
```




tukey: 2.758609 

```{r}
for (l in c(l1, l2, l3, l4, l5, l6)){
    
    print(2 * l / sqrt(s2))
    
}
    
```

significant pairs:

all except $\mu_1 - \mu_2$

Scheffe: 0.5893068 

```{r}
for (l in c(l1, l2, l3, l4, l5, l6)){
    
    print(l^2 / (3 * s2))
    
}
```


all except $\mu_1 - \mu_2$

Bonferroni: 1.456063 

```{r}
for (l in c(l1, l2, l3, l4, l5, l6)){
    
    print(l / sqrt(s2))
    
}
```


all except $\mu_1 - \mu_2, \mu_4 - \mu_3$


## f) 

The benferroni is more conservative compared to other two.
Scheffe is usually used when the sample size are different.
since the sample sizes are equal the tukey method is more appropriate for this situations.