---
title: "HW3"
author: "Sam Mottahedi"
date: "July 16, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(SuppDists)
require(ggplot2)
library(lawstat)
```


# Problem 1

## a)

```{r}
oxy1 = c(1,5,2,1,2,2,4,3,0,2) 
oxy5 = c(4,8,2,3,8,5,6,4,3,3)
oxy10 = c(20,26,24,11,28,20,19,19,21,24) 
oxy20 = c(37,30,26,24,41,25,36,31,31,33)

df <-  data.frame(oxy1, oxy5, oxy10, oxy20)

boxplot(df)
```

No.

## b)


### Hartley test

```{r}
s2 <- apply(df,MARGIN = 2,FUN = var)
max(s2)/min(s2)
qmaxFratio(0.95, 9, 4)
```

$df=9$ , $r=4$


$H(0.95, 4, 9) = 6.31$

$H^* > H$. rejecting the null hypothesis $H_0 = \sigma_1^2 = \dots = \sigma _4^2$

## c)

```{r}
s2 <- apply(sqrt(df),MARGIN = 2,FUN = var)
max(s2)/min(s2)
qmaxFratio(0.95, 9, 4)
```


$df=9$ , $r=4$


$H(0.95, 4, 9) = 6.31$

$H^* \le H$. can't rejecting the null hypothesis $H_0 = \sigma_1^2 = \dots = \sigma _4^2$.


## d) 

```{r}
df <- data.frame(y=c(oxy1, oxy5, oxy10, oxy20), x=as.factor(rep(1:4, each=10)))

fit <- aov(sqrt(y) ~ x, df)

qqnorm(fit$residuals)
```

The residual plots does not indicate serious departure from normality except the two point at the bottom tail.



# Problem 2

## a)

```{r}
mu <- matrix(c(10, 5 ,9, 8, 12 ,7 ,11 ,10, 8 ,3 ,7 ,6), nrow=3, ncol=4, byrow=T)

mudot <- sum(mu)/(nrow(mu) * ncol(mu))

```


$\mu_{..} = `r mudot`$

$\mu_{.j} = `r apply(mu, 2, sum)/3`$

$\mu_{i.} = `r apply(mu, 1, sum)/4`$


$\alpha_{1,2,3} = `r apply(mu, 1, sum)/4 - mudot`$

$\beta_{1,2,3,4} = `r apply(mu, 2, sum)/3 - mudot`$


## b)

because it is possible to express all $\mu_{ij}$ as the sum of $\mu_{..} + \alpha_i + \beta_k$. there's no interaction between the two factors of a and b.



# Problem 3

## a)

```{r}
mu <- matrix(c(8, 5, 9 ,10 ,12 ,11, 7 ,10, 8, 3 ,7 ,6), nrow=3, ncol=4, byrow=T)

mudot <- sum(mu)/(nrow(mu) * ncol(mu))

```


$\mu_{..} = `r mudot`$

$\mu_{.j} = `r apply(mu, 2, sum)/3`$

$\mu_{i.} = `r apply(mu, 1, sum)/4`$


$\alpha_{1,2,3} = `r apply(mu, 1, sum)/4 - mudot`$

$\beta_{1,2,3,4} = `r apply(mu, 2, sum)/3 - mudot`$


## b)

because it's not possible to express all $\mu_{ij}$ as the sum of $\mu_{..} + \alpha_i + \beta_k$. there are interaction between the two factor of a and b.



# Problem 4

## a)

```{r}
df <- read.table('./CH19PR12.txt')
df$V2 <- as.factor(df$V2)
df$V3 <- as.factor(df$V3)
interaction.plot(df$V2, df$V3, df$V1)
```

since the slope of the lines are not zero it is the evidence that V2 has an effect and since the the two lines indicating the levels ov V3 have different heights the V3 has an effect. The two lines are almost parrallel to each other which shows that there no significant interaction.


## b)

```{r}
fit <- lm(V1 ~ V2 + V3 + V2 * V3 , df)
anova(fit)
```

## c)


$H_0: all (\alpha \beta)_{ij} = 0$

$H_a : not all (\alpha \beta)_{ij}$ equal zero


$F^* = \frac{MSAB}{MSE} = `r 1.250 / 6.075`$

$F=`r qf(0.95, 1, 36)`$

$F^* < F$ concluding $H_0$, eye contact and sex do not interact in their effect on applicant job sucess.

## d)

for eye contact:

$H_0: \alpha_1 = \dots = \alpha_2 = 0$

$H_a: \textrm(not all) \alpha_i \textrm{equal zero}$


$F^* = \frac{54.450}{6.075} = `r 54.450 / 6.075`$

$F = `r qf(0.95, 1, 9*4)`$

$F^* > F$, we conclude $H_a$ that factor eye contact means are not equal and  there's some definite effect associated with eye contact and level of sucess.


for sex:

$H_0: \beta_1 = \dots = \beta_2 = 0$

$H_a: \textrm(not all) \beta_i \textrm{equal zero}$

$F^* = \frac{76.050}{6.075} = `r 76.050 / 6.075`$

$F = `r qf(0.95, 1, 9*4)`$


$F^* > F$, we conclude $H_a$ that factor sex means are not equal and  there's some definite effect associated with sex and level of sucess.